---
title: "Create SWMPr Object from non-CDMO data"
author: "Dave Eslinger"
date: "10/27/2020, revised 9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, include=FALSE}
suppressPackageStartupMessages({
library(SWMPrExtension)
library(tidyverse)
library(broom)
})
```

## Assumptions

We have a large amount of monitoring data and want to take advantage of the SWMPr and SWMPrExtension packages, however, the data is not actually SWMP data, so doens't come in the usual format that one gets SWMP data in from CDMO.

## With Non-SWMP data

First load the raw non-SWMP CSV data

```{r read_csv}
raw_csv <- read.csv(file = "./BBAP_BB14_Master_data_11.6.2020.csv")
str(raw_csv)
```

The `swmpr()` call needs to have just datetimestamp and data+qa columns, so remove the extras, while also making names lower case.  

```{r remove_non-SWMP_variables}
names(raw_csv) <- tolower(names(raw_csv))
# names(raw_csv) 
# Just did this to check all the names.  Get rid of all but
# datetimestamp and data + f_data columns.

raw2 <- select(raw_csv, -c(station.code, f_record, rowid, bgaperfu, bgapeug.l, condms.cm, nlfcondms.cm, odo.local, pressurepsia, tdsmg.l, phmv))
# raw2 <- select(raw_csv, -c(stationcode, f_record))
```

A swmpr object needs a time zone 
```{r fix_timestamp}
timezone <- "America/Jamaica"  #Default to EST.  Note that SWMPr does not use daylight savings time.

raw2$datetimestamp <- as.POSIXct(raw2$datetimestamp, tz = timezone, format = '%m/%d/%Y %H:%M')    

```

and a "reserve name", which can be 5 characters (usually 3-letter reserve abbreviation and 2 char. site abbreviation) and a 2-3 letter abbreviation for the data type, one of "wq", "nut", or "met".  I'm not sure this matters.

```{r create_swmpr_object}
swmp1 <- swmpr(raw2,"abc12wq")

class(swmp1)
str(swmp1)
#head(swmp1)
```

Seems to work, or at least didn't break.  Let's try some QA

```{r do_qaqc}
# First check the QAQC flags:

flags <- qaqcchk(swmp1)  # Make a table of all the flags

flags[flags$flag == "0", ]  # check number & type of"Good" data, flag = 0

swmp2 <- tail(swmp1) %>% 
  select(starts_with("f_")) #  Just looking at the last flags here.  Not sure why I did that, actually.

swmp1qc <- qaqc(swmp1, qaqc_keep = "0")  #Keep just the good data.
```

And some plotting
```{r make_test_plots}
seasonal_boxplot(swmp1qc, param = 'do_mgl', target_yr = '2019')
threshold_criteria_plot(WQ, param = 'do_mgl', rng = 2020, thresholds = c(2,5))
```

That thrshold criteria plot looks odd, plot just the do_mgl outside of SWMPrExtension to check
```{r check_plotting}
plot(swmp1qc$datetimestamp, swmp1qc$do_mgl)
```

OK, so it appears ot just be odd-looking data.
