---
title: "Hurricane Ian"
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  error: false
filters: 
  - lightbox
lightbox: auto
---

```{r}
#| label: data-load
#| include: false
library(here)

source(here::here('R', '00_loadpackages.R'))
source(here::here('R', '02.1_load_wrangle_NUT.R'))
source(here::here('R', '02.2_load_wrangle_WQ-and-MET.R'))
source(here::here('R', '03_graph_fxns.R'))
```

![Hurricane Ian's full journey across the Caribbean, through the Gulf of Mexico, and into the Southeast U.S. 09/23/2022 00:00 UTC- 10/01/2022 00:00 UTC. GeoColor product from NOAA GOES-16 satellite. Accessed: https://satlib.cira.colostate.edu/event/hurricane-ian/](/data/hurricane-ian/202209230000-202210010010_g16_fd_geocolor_ianfulljourney_nolabels.mp4){#fig-ian}

Hurricane Ian (@fig-ian) made landfall near Cayo Costa in southwestern Florida at 3:05pm EDT on Wednesday, September 28, 2022 as a Category 4 storm. It then made it's way across the Florida peninsula to Cape Canaveral while downgrading to a Tropical Storm, before hitting the Atlantic Ocean in the early evening of Thursday, September 29, 2022 where it then strengthened to a Category 1 storm and headed towards the South Carolina coast. It made landfall south of Georgetown, South Carolina as a Category 1 storm the afternoon of Friday, September 30, 2022.

The effects of Hurricane Ian were observable at the GTM Research Reserve in northeast Florida starting Wednesday, September 28, 2022 with the onset of northeasterly winds and rising tides. The storm was closest to the GTM Research Reserve at 2:30 EST as determined by the lowest barometric pressure recorded at the weather station in Pellicer Creek (1001 mb).

## Onset of winds

The winds picked up in northeast Florida the morning of Wednesday, September 28, 2022 from the northeast direction. News reports of the region [experiencing a Nor'easter](https://floridapolitics.com/archives/559346-duval-ian/) began on September 28th and the weather station in Pellicer Creek supports the onset of predominating northeasterly winds midday on Tuesday, September 27th that extended until after the storm went off the coast the evening of Thursday, September 29th (@fig-winds). The station experienced sustained winds in excess of 30mph from 5am until almost 5pm on Thursday, September 29th when the storm moved offshore (@fig-winds-1). Maximum gusts during this same period of time were greater than 45 mph with the greatest gusts recorded at 48 mph (@fig-winds-2).

```{r}
#| label: fig-winds
#| echo: false
#| fig.cap: Hourly wind speed and direction from the Pellicer Creek weather station during timeperiod of Hurricane Ian within the state of Florida, USA (Tuesday, September 27 - Saturday, October 1, 2022). The warmer the color, the higher the recorded wind speed. The arrows indicate the direction the wind was blowing towards and the length of the arrow reflects the wind speed information with longer meaning faster. All data is based upon 5 second readings at the weather station.
#| fig.subcap:
#|    - "Averaged hourly wind speeds"
#|    - "Maximum wind gust recorded within each hour"

# helpful resources
# https://stackoverflow.com/questions/47880918/how-to-plot-wind-direction-with-lat-lon-and-arrow-in-ggplot2 
# https://stackoverflow.com/questions/49639881/how-to-get-a-legend-on-geom-spoke

# create new df for just wspd and direction data with conversions to SI units
MET_wind <- MET %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-09-27 00:00:00"), 
                 as.POSIXct("2022-10-02 00:00:00"))) %>%
  aggreswmp(by = "hours", params = c('wspd', 'maxwspd', 'wdir')) %>% 
  mutate(day = lubridate::date(datetimestamp),
         hour = lubridate::hour(datetimestamp),
         wspd_mph = wspd * 2.23694,
         maxwspd_mph = maxwspd * 2.23694,
         wdir_x = ((wdir +180) * 0.0174533)) # angles are in radians not degrees, need to convert for plotting


MET_wind %>% 
  filter(day > as.Date("2022-09-26")) %>% 
  ggplot(aes(x = hour, 
             y = day, 
             fill = wspd_mph, 
             angle = wdir_x, 
             radius = scales::rescale(wspd_mph, c(.3, .8)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlBu") + 
    coord_equal(expand = 0) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14,16,18,20,22)) +
    theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        legend.position = 'top', 
        legend.direction = 'horizontal') +
  labs(x = "Hour of Day (24hr)",
       y = "",
       fill = "Wind Speed (mph)")


MET_wind %>% 
  filter(day > as.Date("2022-09-26")) %>% 
  ggplot(aes(x = hour, 
             y = day, 
             fill = maxwspd_mph, 
             angle = wdir_x, 
             radius = scales::rescale(wspd_mph, c(.3, .8)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlBu") + 
    coord_equal(expand = 0) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14,16,18,20,22)) +
    theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        legend.position = 'top', 
        legend.direction = 'horizontal',
        axis.line = element_blank()) +
  labs(x = "Hour of Day (24hr)",
       y = "",
       fill = "Max Wind Gusts (mph)")

```

## Rising tides

Water levels began to rise in the St. Augustine area Wednesday, September 28, 2022. Unlike storms Matthew and Irma in the past (@fig-hist-storms), there was no observable singular storm surge event that occurred with Hurricane Ian, but increasing water levels over several tidal cycles (@fig-surge). It took three tide cycles until maximum water levels were reached when the storm was closest to the reserve. For most stations, the amount of water at the incoming tide was close to double what was receding at the outgoing tide. The greatest increase in water levels were observed in Pellicer Creek where there was an overall gain of 1.31m (4.3ft) in the creek. After the storm's passing the water levels quickly dropped with each falling tide over the next two days (@fig-surge).

```{r}
#| label: fig-hist-storms
#| echo: false
#| fig.cap: Continuous water depth at all four water quality stations of the GTMNERR during the time period of Hurricanes Matthew, October 6-9, 2016 and Irma, September 9-12, 2017.

irma <- WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2017-09-09 00:00:00"), 
                 as.POSIXct("2017-09-12 00:00:00"))) %>%
  mutate(station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1,
         storm = "Irma (2017)") %>% 
  select(station, datetimestamp, depth, storm)  
matt <- WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2016-10-06 00:00:00"), 
                 as.POSIXct("2016-10-09 00:00:00"))) %>%
  mutate(station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1,
         storm = "Matthew (2016)") %>% 
  select(station, datetimestamp, depth, storm)

hist_storms <- bind_rows(irma, matt) %>% mutate(storm = factor(storm,
                                                               levels = c("Matthew (2016)",
                                                                          "Irma (2017)"),
                                                               ordered = T))

hist_storms %>% 
  ggplot(aes(x = datetimestamp, y = depth, color = station, group = station)) +
  geom_line(size = 1) +
  facet_grid(~storm, scales = "free_x") +
  scale_color_okabeito(breaks = c("Pine Island", "San Sebastian", "Fort Matanzas", "Pellicer Creek")) +
  scale_y_continuous(limits = c(0,9),
                     breaks = seq(0, 8, by = 2),
                     minor_breaks = seq(0,9, by = 1),
                     expand = c(0,0)) +
  scale_x_datetime(date_breaks = "1 day",
                   date_minor_breaks = "6 hours",
                   date_labels = "%b %d") +
  # theme_classic() +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray", linetype = "dotted"),
        legend.key = element_rect(fill = NA),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")) +
  labs(x = "",
       y = "Water Depth (m)",
       color = "SWMP Station") 

rm(irma, matt, hist_storms)
```


```{r}
#| label: fig-surge
#| echo: false
#| fig.cap: Continuous water depth at all four water quality stations of the GTMNERR during the time period of Hurricane Ian, September 28 - October 2, 2022. Red line indicates when the Hurricane was closest to the GTMNERR weather station in Pellicer Creek, based on barometric pressure readings.

library(gganimate)

d_animation <-
WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-09-28 00:00:00"), 
                 as.POSIXct("2022-10-02 00:00:00"))) %>%
  mutate(depth = if_else(station == "gtmpcwq", 
                         level + 1.11427, 
                         depth),
         station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1) %>% # convert PC level to depth to be able to compare sites
  select(station, datetimestamp, depth) %>% 
  ggplot(aes(x = datetimestamp, y = depth, color = station, group = station)) +
  geom_vline(xintercept = as.POSIXct("2022-09-29 14:30:00"), color = "red", size = 1) +
  geom_line() +
  geom_point(size = 3) +
  scale_color_okabeito(breaks = c("Pine Island", "San Sebastian", "Fort Matanzas", "Pellicer Creek")) +
  scale_y_continuous(limits = c(0,9),
                     breaks = seq(0, 8, by = 2),
                     minor_breaks = seq(0,9, by = 1),
                     expand = c(0,0)) +
  scale_x_datetime(date_breaks = "1 day",
                   date_minor_breaks = "6 hours",
                   date_labels = "%b %d") +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray", linetype = "dotted"),
        legend.key = element_rect(fill = NA),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")) +
  labs(x = "",
       y = "Water Depth (m)",
       color = "SWMP Station") +
  annotate("text",
           x = as.POSIXct("2022-09-29 13:00:00"),
           y = 1.5,
           label = "Ian 2022-09-29 14:30 EST",
           color = "red",
           angle = 90,
           size = 2.1) +
  gganimate::transition_reveal(datetimestamp)

gganimate::animate(d_animation, height = 5, width = 8, units = "in", res = 120)
```

## Heavy rainfall

```{r}

```

## Loaded freshwater system
