---
title: "Hurricane Ian"
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  error: false
link-external-icon: true
link-external-newwindow: true
filters: 
  - lightbox
lightbox: auto
---

```{r}
#| label: data-load
#| include: false
library(here)

source(here::here('R', '00_loadpackages.R'))
source(here::here('R', '02.1_load_wrangle_NUT.R'))
source(here::here('R', '02.2_load_wrangle_WQ-and-MET.R'))
source(here::here('R', '03_graph_fxns.R'))
```

![Hurricane Ian's full journey across the Caribbean, through the Gulf of Mexico, and into the Southeast U.S. 09/23/2022 00:00 UTC- 10/01/2022 00:00 UTC. GeoColor product from NOAA GOES-16 satellite. Accessed: https://satlib.cira.colostate.edu/event/hurricane-ian/](/data/hurricane-ian/202209230000-202210010010_g16_fd_geocolor_ianfulljourney_nolabels.mp4){#fig-ian}

Hurricane Ian (@fig-ian) made landfall near Cayo Costa in southwestern Florida at 3:05pm EDT on Wednesday, September 28, 2022 as a Category 4 storm. It then made it's way across the Florida peninsula to Cape Canaveral while downgrading to a Tropical Storm, before hitting the Atlantic Ocean in the early evening of Thursday, September 29, 2022 where it then strengthened to a Category 1 storm and headed towards the South Carolina coast. It made landfall south of Georgetown, South Carolina as a Category 1 storm the afternoon of Friday, September 30, 2022.

The effects of Hurricane Ian were observable at the GTM Research Reserve in northeast Florida starting Wednesday, September 28, 2022 with the onset of northeasterly winds and rising tides. The storm was closest to the GTM Research Reserve at 2:30 EST as determined by the lowest barometric pressure recorded at the weather station in Pellicer Creek (1001 mb).

## Onset of winds

The winds picked up in northeast Florida the morning of Wednesday, September 28, 2022 from the northeast direction. News reports of the region [experiencing a Nor'easter](https://floridapolitics.com/archives/559346-duval-ian/) began on September 28th and the weather station in Pellicer Creek supports the onset of predominating northeasterly winds midday on Tuesday, September 27th that extended until after the storm went off the coast the evening of Thursday, September 29th (@fig-winds). The station experienced sustained winds in excess of 30mph from 5am until almost 5pm on Thursday, September 29th when the storm moved offshore (@fig-winds-1). Maximum gusts during this same period of time were greater than 45 mph with the greatest gusts recorded at 48 mph (@fig-winds-2).

```{r}
#| label: fig-winds
#| echo: false
#| fig.cap: Hourly wind speed and direction from the Pellicer Creek weather station during timeperiod of Hurricane Ian within the state of Florida, USA (Tuesday, September 27 - Saturday, October 1, 2022). The warmer the color, the higher the recorded wind speed. The arrows indicate the direction the wind was blowing towards and the length of the arrow reflects the wind speed information with longer meaning faster. All data is based upon 5 second readings at the weather station.
#| fig.subcap:
#|    - "Averaged hourly wind speeds"
#|    - "Maximum wind gust recorded within each hour"
#| layout-ncol: 2

# helpful resources
# https://stackoverflow.com/questions/47880918/how-to-plot-wind-direction-with-lat-lon-and-arrow-in-ggplot2 
# https://stackoverflow.com/questions/49639881/how-to-get-a-legend-on-geom-spoke

# create new df for just wspd and direction data with conversions to SI units
MET_wind <- MET %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-09-27 00:00:00"), 
                 as.POSIXct("2022-10-02 00:00:00"))) %>%
  aggreswmp(by = "hours", params = c('wspd', 'maxwspd', 'wdir')) %>% 
  mutate(day = lubridate::date(datetimestamp),
         hour = lubridate::hour(datetimestamp),
         wspd_mph = wspd * 2.23694,
         maxwspd_mph = maxwspd * 2.23694,
         wdir_x = ((wdir +180) * 0.0174533)) # angles are in radians not degrees, need to convert for plotting


MET_wind %>% 
  filter(day > as.Date("2022-09-26")) %>% 
  ggplot(aes(x = hour, 
             y = day, 
             fill = wspd_mph, 
             angle = wdir_x, 
             radius = scales::rescale(wspd_mph, c(.3, .8)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlBu") + 
    coord_equal(expand = 0) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14,16,18,20,22)) +
    theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        legend.position = 'top', 
        legend.direction = 'horizontal') +
  labs(x = "Hour of Day (24hr)",
       y = "",
       fill = "Wind Speed (mph)")


MET_wind %>% 
  filter(day > as.Date("2022-09-26")) %>% 
  ggplot(aes(x = hour, 
             y = day, 
             fill = maxwspd_mph, 
             angle = wdir_x, 
             radius = scales::rescale(wspd_mph, c(.3, .8)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlBu") + 
    coord_equal(expand = 0) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14,16,18,20,22)) +
    theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        legend.position = 'top', 
        legend.direction = 'horizontal',
        axis.line = element_blank()) +
  labs(x = "Hour of Day (24hr)",
       y = "",
       fill = "Max Wind Gusts (mph)")

```

## Rising tides

Water levels began to rise in the St. Augustine area Wednesday, September 28, 2022. Unlike storms Matthew and Irma in the past (@fig-hist-storms), there was no observable singular storm surge event that occurred with Hurricane Ian, but increasing water levels over several tidal cycles (@fig-surge). It took three tide cycles until maximum water levels were reached when the storm was closest to the reserve. For most stations, the amount of water at the incoming tide was close to double what was receding at the outgoing tide. The greatest increase in water depth was observed in Pellicer Creek where there was an overall gain of 1.31m (4.3ft) in the creek. After the storm's passing, the water levels quickly dropped with each falling tide over the next two days (@fig-surge).

```{r}
#| label: fig-hist-storms
#| echo: false
#| fig.cap: Continuous water depth at all four water quality stations of the GTMNERR during the time periods of Hurricanes Matthew, October 6-9, 2016 and Irma, September 9-12, 2017 showing the storm surge within one tidal cycle that occurred during both storms.

irma <- WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2017-09-09 00:00:00"), 
                 as.POSIXct("2017-09-12 00:00:00"))) %>%
  mutate(station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1,
         storm = "Irma (2017)") %>% 
  select(station, datetimestamp, depth, storm)  
matt <- WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2016-10-06 00:00:00"), 
                 as.POSIXct("2016-10-09 00:00:00"))) %>%
  mutate(station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1,
         storm = "Matthew (2016)") %>% 
  select(station, datetimestamp, depth, storm)

hist_storms <- bind_rows(irma, matt) %>% mutate(storm = factor(storm,
                                                               levels = c("Matthew (2016)",
                                                                          "Irma (2017)"),
                                                               ordered = T))

hist_storms %>% 
  ggplot(aes(x = datetimestamp, y = depth, color = station, group = station)) +
  geom_line(size = 1) +
  facet_grid(~storm, scales = "free_x") +
  scale_color_okabeito(breaks = c("Pine Island", "San Sebastian", "Fort Matanzas", "Pellicer Creek")) +
  scale_y_continuous(limits = c(0,9),
                     breaks = seq(0, 8, by = 2),
                     minor_breaks = seq(0,9, by = 1),
                     expand = c(0,0)) +
  scale_x_datetime(date_breaks = "1 day",
                   date_minor_breaks = "6 hours",
                   date_labels = "%b %d") +
  # theme_classic() +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray", linetype = "dotted"),
        legend.key = element_rect(fill = NA),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")) +
  labs(x = "",
       y = "Water Depth (m)",
       color = "SWMP Station") 

rm(irma, matt, hist_storms)
```

```{r}
#| label: fig-surge
#| echo: false
#| fig.cap: Continuous water depth at all four water quality stations of the GTMNERR during the time period of Hurricane Ian, September 28 - October 2, 2022. Red line indicates when the Hurricane was closest to the GTMNERR weather station in Pellicer Creek, based on barometric pressure readings.

d_animation <-
WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-09-28 00:00:00"), 
                 as.POSIXct("2022-10-02 00:00:00"))) %>%
  mutate(depth = if_else(station == "gtmpcwq", 
                         level + 1.11427, 
                         depth),
         station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1) %>% # convert PC level to depth to be able to compare sites
  select(station, datetimestamp, depth) %>% 
  ggplot(aes(x = datetimestamp, y = depth, color = station, group = station)) +
  geom_vline(xintercept = as.POSIXct("2022-09-29 14:30:00"), color = "red", size = 1) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  scale_color_okabeito(breaks = c("Pine Island", "San Sebastian", "Fort Matanzas", "Pellicer Creek")) +
  scale_y_continuous(limits = c(0,9),
                     breaks = seq(0, 8, by = 2),
                     minor_breaks = seq(0,9, by = 1),
                     expand = c(0,0)) +
  scale_x_datetime(date_breaks = "1 day",
                   date_minor_breaks = "6 hours",
                   date_labels = "%b %d") +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray", linetype = "dotted"),
        legend.key = element_rect(fill = NA),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")) +
  labs(x = "",
       y = "Water Depth (m)",
       color = "SWMP Station") +
  annotate("text",
           x = as.POSIXct("2022-09-29 13:00:00"),
           y = 1.5,
           label = "Ian 2022-09-29 14:30 EST",
           color = "red",
           angle = 90,
           size = 2.1) +
  gganimate::transition_reveal(datetimestamp)

gganimate::animate(d_animation, 
                   height = 5, width = 8, units = "in", 
                   res = 120,
                   end_pause = 50)

rm(d_animation)
```

## Heavy rainfall

All the rainfall during the storm occurred within two days Wednesday, September 28 and Thursday September 29, 2022. After the storm passed into the Atlantic Ocean the night of September 29, there was an abrupt stop of all precipitation and the system did not get any more rain until more than two weeks later on October 17, 2022. There was a total of 238mm (9.37in) of rain within those two days, the majority of which (67.5%) was on Thursday, September 29, 2022. (*All information based on data collected at the weather station in Pellicer Creek*).

## A loaded freshwater system

Between the increasing water levels before the storm and the heavy rainfall with it, the salinity at all the water quality stations was quite variable (@fig-salinity). For the two stations furthest away from the inlets (Pine Island and Pellicer Creek), salinity started to increase with each tidal cycle starting on Wednesday, September 28th, 2022. The predominating northeasterly winds along the northeast Florida coast at this time largely fought falling tides, trapping more salty, ocean waters within the estuary.

All sites experienced drops in salinity as the storm passed, but the greatest effect was in Pellicer Creek where salinity dropped to 0psu (fresh water) for four days before any observable rise in salinity was observed by the station. Quite a bit of freshwater movement and draining from the watershed is observable at the other three stations in the days after the storm (@fig-salinity). Large changes (peaks and drops) in salinity at Fort Matanzas is most likely due to the influence of Pellicer Creek, the largest freshwater source to the Matanzas Estuary, as well as the proximity of this site to the Matanzas Inlet (4 km north). The Pine Island station appeared to have increased in salinity with the storm and then fell back to pre-storm levels within a day and a half after the storm's passing.

```{r}
#| label: fig-salinity
#| echo: false
#| fig.cap: Continuous salinity at all four water quality stations of the GTMNERR during the time period of Hurricane Ian, September 28 - October 2, 2022. Red line indicates when the Hurricane was closest to the GTMNERR weather station in Pellicer Creek, based on barometric pressure readings.

sal_animation <-
WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-09-27 00:00:00"), 
                 as.POSIXct("2022-10-03 00:00:00"))) %>%
  mutate(station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F)) %>% 
  select(station, datetimestamp, sal) %>% 
  ggplot(aes(x = datetimestamp, y = sal, color = station, group = station)) +
  geom_vline(xintercept = as.POSIXct("2022-09-29 14:30:00"), color = "red", size = 1) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  scale_color_okabeito(breaks = c("Pine Island", "San Sebastian", "Fort Matanzas", "Pellicer Creek")) +
  scale_x_datetime(date_breaks = "1 day",
                   date_minor_breaks = "6 hours",
                   date_labels = "%b %d") +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray", linetype = "dotted"),
        legend.key = element_rect(fill = NA),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")) +
  labs(x = "",
       y = "Salinity (psu)",
       color = "SWMP Station") +
  annotate("text",
           x = as.POSIXct("2022-09-29 13:00:00"),
           y = 5,
           label = "Ian 2022-09-29 14:30 EST",
           color = "red",
           angle = 90,
           size = 2.1) +
  gganimate::transition_reveal(datetimestamp)


gganimate::animate(sal_animation, 
                   height = 5, width = 8, units = "in", 
                   res = 120,
                   end_pause = 50)

rm(sal_animation)
```

## Recovery

A large note to make when investigating recovery time frames is that not long after the passing of the hurricane, northeast Florida experienced seasonal "king tide" events, which regularly occur in the fall months (primarily October) and lead to higher than high tides. This effect largely impacts the ability to fully assess the "recovery" of the system after the hurricane. However, observing potential recovery within two weeks after the storm revealed that many sites did return to pre-storm values or ranges (@tbl-days).

Dissolved oxygen at many stations increased with the storm and remained high for several days following the storm. Preceding the storm, much of the estuary was experiencing lower values of dissolved oxygen. The storm appeared to have reoxygenated the waterways, though both Fort Matanzas and San Sebastian experienced a drop in the day after the storm, which then climbed back up to pre-storm levels within a few days. For Pine Island, the dissolved oxygen remained high after the storm and hasn't fallen back to pre-storm patterns.

Although pre-storm salinity values were reached not long after the storm passed at San Sebastian and Fort Matanzas, the tidal range of salinity did not resume pre-storm patterns until over a week had passed after the storm. This was likely largely driven by the excess freshwater in the estuary.

```{r}
#| label: tbl-days
#| echo: false
#| tbl-cap: Number of days it took to return to pre-storm values/patterns after the passing of Hurricane Ian at all four water quality stations in the GTM Research Reserve.


days <-
  tribble(
  ~station, ~salinity, ~do, ~depth,
  "Pine Island", 1, NA, 1,
  "San Sebastian", 11, 6, 1,
  "Fort Matanzas", 12, 4, 1,
  "Pellicer Creek", 12, 10, 13
)

days %>% 
  rename(Station = station,
         Salinity = salinity,
         `Dissolved Oxygen` = do,
         Depth = depth) %>% 
  kbl(align = "c") %>% 
  kable_styling(fixed_thead = T) %>% 
  row_spec(0, bold = T) %>% 
  column_spec(1, border_right = T) %>%
  kable_classic(c('striped', 'hover'), full_width = F)

```
