---
title: "Hurricane Nicole"
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  error: false
link-external-icon: true
link-external-newwindow: true
filters: 
  - lightbox
lightbox: auto
---

```{r}
#| label: data-load
#| include: false
library(here)

source(here::here('R', '00_loadpackages.R'))
source(here::here('R', '02.1_load_wrangle_NUT.R'))
source(here::here('R', '02.2_load_wrangle_WQ-and-MET.R'))
source(here::here('R', '03_graph_fxns.R'))
```

![Hurricane Nicole's swept across the peninsula of Florida, USA, 11/10/2022 00:00 UTC- 11/11/2022 17:00 UTC. GeoColor product from NOAA GOES-16 satellite. Accessed: https://satlib.cira.colostate.edu/event/nicole/](/data/hurricane-nicole/202211100001-202211111711_g16_conus_geocolor_nicole_labels.mp4){#fig-nicole}

Hurricane Nicole formed on the morning of Monday, November 7th, 2022 and became a tropical storm the following morning on Tuesday, November 8th, 2022 before strengthening to a Category 1 Hurricane on Wednesday, November 9th, 2022. It made landfall near Vero Beach, Florida, USA on Thursday, November 10th and dissipated across the New England part of the US on Friday, November 11th, 2022. 

## Onset of Winds

```{r}
#| label: fig-winds-nicole
#| echo: false
#| fig.cap: Hourly wind speed and direction from the Pellicer Creek weather station during timeperiod of Hurricane Nicole within the state of Florida, USA (Wednesday, November 9 - Friday, November 11, 2022). The warmer the color, the higher the recorded wind speed. The arrows indicate the direction the wind was blowing towards and the length of the arrow reflects the wind speed information with longer meaning faster. All data is based upon 5 second readings at the weather station.
#| fig.subcap:
#|    - "Averaged hourly wind speeds"
#|    - "Maximum wind gust recorded within each hour"
#| layout-ncol: 2

# helpful resources
# https://stackoverflow.com/questions/47880918/how-to-plot-wind-direction-with-lat-lon-and-arrow-in-ggplot2 
# https://stackoverflow.com/questions/49639881/how-to-get-a-legend-on-geom-spoke

# create new df for just wspd and direction data with conversions to SI units
MET_wind <- MET %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-11-08 00:00:00"), 
                 as.POSIXct("2022-11-12 00:00:00"))) %>%
  aggreswmp(by = "hours", params = c('wspd', 'maxwspd', 'wdir')) %>% 
  mutate(day = lubridate::date(datetimestamp),
         hour = lubridate::hour(datetimestamp),
         wspd_mph = wspd * 2.23694,
         maxwspd_mph = maxwspd * 2.23694,
         wdir_x = ((wdir +180) * 0.0174533)) # angles are in radians not degrees, need to convert for plotting


MET_wind %>% 
  filter(day > as.Date("2022-11-08")) %>% 
  ggplot(aes(x = hour, 
             y = day, 
             fill = wspd_mph, 
             angle = wdir_x, 
             radius = scales::rescale(wspd_mph, c(.3, .8)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlBu") + 
    coord_equal(expand = 0) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14,16,18,20,22)) +
    theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        legend.position = 'top', 
        legend.direction = 'horizontal') +
  labs(x = "Hour of Day (24hr)",
       y = "",
       fill = "Wind Speed (mph)")


MET_wind %>% 
  filter(day > as.Date("2022-11-08")) %>% 
  ggplot(aes(x = hour, 
             y = day, 
             fill = maxwspd_mph, 
             angle = wdir_x, 
             radius = scales::rescale(wspd_mph, c(.3, .8)))) +
    geom_raster() +
    geom_spoke(arrow = arrow(length = unit(.05, 'inches'))) + 
    scale_fill_distiller(palette = "RdYlBu") + 
    coord_equal(expand = 0) +
    scale_x_continuous(breaks = c(0,2,4,6,8,10,12,14,16,18,20,22)) +
    theme_classic() +
  theme(axis.text = element_text(color = "black", size = 12),
        legend.position = 'top', 
        legend.direction = 'horizontal',
        axis.line = element_blank()) +
  labs(x = "Hour of Day (24hr)",
       y = "",
       fill = "Max Wind Gusts (mph)")

```

## Rising Tides

```{r}
#| label: fig-surge-nicole
#| echo: false
#| fig.cap: Continuous water depth at all four water quality stations of the GTMNERR during the time period of Hurricane Nicole, Tuesday, November 8 - Friday, November 11, 2022. Red line indicates when the Hurricane was closest to the GTMNERR weather station in Pellicer Creek, based on barometric pressure readings.

d_animation <-
WQ %>% 
  filter(between(datetimestamp, 
                 as.POSIXct("2022-11-08 00:00:00"), 
                 as.POSIXct("2022-11-11 00:00:00"))) %>%
  mutate(depth = if_else(station == "gtmpcwq", 
                         level + 1.11427, 
                         depth),
         station = factor(station,
                          levels = c("gtmsswq",
                                     "gtmpiwq",
                                     "gtmfmwq",
                                     "gtmpcwq"),
                          labels = c("San Sebastian",
                                     "Pine Island",
                                     "Fort Matanzas",
                                     "Pellicer Creek"),
                          ordered = F),
         depth = depth + 1) %>% # convert PC level to depth to be able to compare sites
  select(station, datetimestamp, depth) %>% 
  ggplot(aes(x = datetimestamp, y = depth, color = station, group = station)) +
  geom_vline(xintercept = as.POSIXct("2022-11-10 07:30:00"), color = "red", size = 1) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  scale_color_okabeito(breaks = c("Pine Island", "San Sebastian", "Fort Matanzas", "Pellicer Creek")) +
  scale_y_continuous(limits = c(0,9),
                     breaks = seq(0, 8, by = 2),
                     minor_breaks = seq(0,9, by = 1),
                     expand = c(0,0)) +
  scale_x_datetime(date_breaks = "1 day",
                   date_minor_breaks = "6 hours",
                   date_labels = "%b %d") +
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "gray"),
        panel.grid.minor = element_line(color = "gray", linetype = "dotted"),
        legend.key = element_rect(fill = NA),
        axis.title = element_text(face = "bold"),
        legend.title = element_text(face = "bold")) +
  labs(x = "",
       y = "Water Depth (m)",
       color = "SWMP Station") +
  annotate("text",
           x = as.POSIXct("2022-11-10 06:00:00"),
           y = 1.5,
           label = "Nicole 2022-11-10 07:30 EST",
           color = "red",
           angle = 90,
           size = 2.1) +
  gganimate::transition_reveal(datetimestamp)

gganimate::animate(d_animation, 
                   height = 5, width = 8, units = "in", 
                   res = 120,
                   end_pause = 50)

rm(d_animation)
```

## Damage to Florida State Road A1A

The section of Florida State Road A1A that runs through St. Johns and Flagler Counties is an officially designated American Byway recognized as a [National Scenic Byway in 2002](https://scenica1a.org/). Both of the offices for the GTM Research Reserve lie along this route in Ponte Vedra Beach (505 Guana River Road, Ponte Vedra Beach, FL 32082) and in Marineland (9741 Ocean Shore Blvd., St Augustine, FL 32137). 
During Hurricane Nicole, predominanting northeast and easterly winds, coupled with increased wave action and high tides, led to the wash out of several sections of state road A1A near both the northern and southern reserve offices. 