---
title: "Other Indicators"
execute: 
  warning: false
  error: false
link-external-icon: true
link-external-newwindow: true
filters: 
  - lightbox
lightbox: auto
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: met-data-load
#| include: false
library(here)

source(here::here('R', '00_loadpackages.R'))
source(here::here('R', '02.1_load_wrangle_NUT.R'))
source(here::here('R', '03_graph_fxns.R'))
```

## Dissolved Organic Carbon

The highest values in dissolved organic carbon (DOC) are observed at Pellicer Creek, particularly in April of this year, which was the highest reported DOC in the entire SWMP dataset for all combined sites (44.5 mg/L) (@fig-doc-boxplots-1). Elevated DOC levels in April were not unique to Pellicer Creek, as this was a pattern in all the sites (@fig-doc-boxplots). Values dropped in each subsequent month until August, where they rose again. They were then further elevated in October after the passage of Hurricane Ian. 

```{r}
#| label: fig-doc-boxplots
#| echo: false

boxplot_currentyear(station = "gtmpinut", 
                     param = doc, 
                     threshold = FALSE) +
 labs(y = "Mean Dissolved Organic Carbon (mg/L)")

boxplot_currentyear(station = "gtmssnut", 
                     param = doc, 
                     threshold = FALSE) +
 labs(y = "Mean Dissolved Organic Carbon (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = doc, 
                     threshold = FALSE) +
 labs(y = "Mean Dissolved Organic Carbon (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = doc, 
                     threshold = FALSE) +
 labs(y = "Mean Dissolved Organic Carbon (mg/L)")

```

## Bacteria

The SWMP program began analyzing water samples at each SWMP station for fecal coliform bacteria in 2013. In 2019, collections for *Enterococcus* bacteria began and both are now collected at all four sites. 

```{r}
#| label: fig-bacteria
#| echo: false

misc_dat <- NUT %>% 
            select(station_code, date_time_stamp, entero_mpn, feccol_cfu, )

boxplot_currentyear(station = "gtmpcnut", 
                     param = entero_mpn, 
                     threshold = FALSE) +
 labs(y = "Mean Enterococcus (MPN)")

```
