---
title: "Nutrient Indicators"
execute: 
  warning: false
  error: false
link-external-icon: true
link-external-newwindow: true
filters: 
  - lightbox
lightbox: auto
bibliography: references.bib
---

```{r}
#| label: data-load
#| include: false
library(here)

source(here::here('R', '00_loadpackages.R'))
source(here::here('R', '02.1_load_wrangle_NUT.R'))
source(here::here('R', '03_graph_fxns.R'))

load(here('output','data', 'pi_wq.Rdata'))
load(here('output','data', 'ss_wq.Rdata'))
load(here('output','data', 'fm_wq.Rdata'))
load(here('output','data', 'pc_wq.Rdata'))
```

The state thresholds used in the chlorophyll *a*, total nitrogen, and total phosphorus visualizations come from Florida State Surface Water Quality Standards within the Florida Administrative Code.

::: callout-note
# Access the Standards

By visiting [Florida Rules](https://www.flrules.org/gateway/ruleno.asp?id=62-302.530)
:::

## Summary

Waters at all stations were oxygenated following spring rainfall events and Nor'easters (@fig-rainfall-3; @fig-do); however, dissolved oxygen was lower than average from late January through mid-March at Pine Island and Pellicer Creek and again at the end July at Pine Island (@fig-do-1; @fig-do-2). This pattern is likely correlated with the warmer than average temperatures this area experienced during this time (@fig-range-atemp; @fig-range-temp). Pellicer Creek had daily average dissolved oxygen values that fell within normal ranges (@fig-do-4), however continuous data show low dissolved oxygen values in both June and July (@fig-do-pc). Pine Island has experienced very low dissolved oxygen values (less than 2 mg/L) frequently in the past two years, which is uncharacteristic for this station, particularly in the past five years (@fig-do-pi). Tables 5.1 and 5.2 show the hypoxia events and duration (more than 0.5 hrs) at Pine Island and Pellicer Creek for this year.  

Chlorophyll *a* values were higher than average at all stations from January through May (@fig-chla-boxplots), frequently over the state thresholds. Although too soon to make inferences, so far the data shows elevated annual geometric means of chlorophyll *a* at all stations except Fort Matanzas (@fig-chla-agm). 

Interpretations of patterns in total nitrogen (TN) and dissolved inorganic nitrogen (DIN) are difficult this year since quite a bit of the values were not able to be calculated. Total Kjeldahl Nitrogen (TKN) was lower than average for most stations; however, all stations experienced higher levels in May (@fig-tkn-boxplots).

Total phosphorus continues to fall below the state threshold for every station and similar to TKN, all stations experienced elevated values of total phosphorus and orthophosphate in May (@fig-tp-boxplots; @fig-po4-boxplots). Relatively high levels of orthophosphate were observed at Pine Island January through May (@fig-po4-boxplots). 

## Data

### Dissolved Oxygen

```{r}
#| label: fig-do
#| echo: false
#| fig.cap: Historical daily range of dissolved oxygen at each SWMP water quality station with current year daily average overlaid in blue. Horizontal dashed red line indicates 2mg/L hypoxia threshold.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

pi |> 
  filter(do_mgl > 0) |> 
historical_daily_range(param = 'do_mgl', target_yr = 2023, criteria = 2)

historical_daily_range(ss, param = 'do_mgl', target_yr = 2023, criteria = 2)

historical_daily_range(filter(fm, do_mgl > 0), param = 'do_mgl', target_yr = 2023, criteria = 2)

pc |>
  filter(do_mgl < 20 & do_mgl > 0) |> 
  historical_daily_range(param = 'do_mgl', target_yr = 2023, criteria = 2)

```

Pellicer Creek is located within a waterbody with known impairments for dissolved oxygen, chlorophyll *a*, and fecal coliform bacteria. As such, this station frequently experiences low dissolved oxygen values, particularly in the summer months (@fig-do-pc).

```{r}
#| label: fig-do-pc
#| echo: false
#| fig.cap: Continuous dissolved oxygen at Pellicer Creek with monthly average overlaid in solid gray line. Conditions are indicated as good when above 5 mg/L, fair between 2 and 5 mg/L, and poor if less than 2 mg/L.
#| fig.subcap:
#|     - "For the past five years"
#|     - "For the current 2023 year"
#| layout-ncol: 2
    
threshold_criteria_plot(pc, param = 'do_mgl', rng = c(2018, 2023), 
                        thresholds = c(2,5),
                        threshold_labs = c('Poor', 'Fair', 'Good'), 
                        monthly_smooth = TRUE, 
                        threshold_cols = c('#FEC596', '#FFFFCC', '#ABD9E9')) +
  theme(axis.text = element_text(color = "black", size = 18),
        axis.title.y = element_text(size = 20))



threshold_criteria_plot(pc, param = 'do_mgl', rng = c(2023, 2023), 
                        thresholds = c(2,5),
                        threshold_labs = c('Poor', 'Fair', 'Good'), 
                        monthly_smooth = TRUE, 
                        threshold_cols = c('#FEC596', '#FFFFCC', '#ABD9E9')) +
  theme(axis.text = element_text(color = "black", size = 18),
        axis.title.y = element_text(size = 20))

```

```{r}
#| label: pc-do-tbl
#| echo: false

pc %>% 
  filter(datetimestamp > "2023-01-01 00:00") %>% 
  setstep() %>% 
  threshold_identification(param = c('do_mgl'),
                           parameter_threshold = c(2),
                           threshold_type = c('<'),
                           time_threshold = 0.5) %>% 
  select(-parameter, -thr_violation, -statement) %>% 
  rename(Duration = duration,
         `Start Time` = starttime,
         `End Time` = endtime) %>% 
  kbl(align = "c", caption = "Table 5.1 Identification and Duration of Hypoxia Events (< 2 mg/L for > 0.5 hrs) at Pellicer Creek") %>%
  kable_styling(fixed_thead = T) %>% 
  column_spec(3, border_left = T)
  
  
```

Dissolved oxygen at Pine Island has been uncharacteristically low in the past two years.

```{r}
#| label: fig-do-pi
#| echo: false
#| fig-cap: Continuous dissolved oxygen at Pine Island with monthly average overlaid in solid gray line. Conditions are indicated as good when above 5 mg/L, fair between 2 and 5 mg/L, and poor if less than 2 mg/L.
#| fig.subcap:
#|     - "For the past five years"
#|     - "For the current 2023 year"
#| layout-ncol: 2

threshold_criteria_plot(pi, param = 'do_mgl', rng = c(2018, 2023), 
                        thresholds = c(2,5),
                        threshold_labs = c('Poor', 'Fair', 'Good'), 
                        monthly_smooth = TRUE, 
                        threshold_cols = c('#FEC596', '#FFFFCC', '#ABD9E9')) +
  theme(axis.text = element_text(color = "black", size = 18),
        axis.title.y = element_text(size = 20))

threshold_criteria_plot(pi, param = 'do_mgl', rng = c(2023, 2023), 
                        thresholds = c(2,5),
                        threshold_labs = c('Poor', 'Fair', 'Good'), 
                        monthly_smooth = TRUE, 
                        threshold_cols = c('#FEC596', '#FFFFCC', '#ABD9E9')) +
  theme(axis.text = element_text(color = "black", size = 18),
        axis.title.y = element_text(size = 20))

```

```{r}
#| label: pi-do-tbl
#| echo: false

pi %>% 
  filter(datetimestamp > "2023-01-01 00:00") %>% 
  setstep() %>% 
  threshold_identification(param = c('do_mgl'),
                           parameter_threshold = c(2),
                           threshold_type = c('<'),
                           time_threshold = 0.5) %>% 
  select(-parameter, -thr_violation, -statement) %>% 
  rename(Duration = duration,
         `Start Time` = starttime,
         `End Time` = endtime) %>% 
  kbl(align = "c", caption = "Table 5.2 Identification and Duration of Hypoxia Events (< 2 mg/L for > 0.5 hrs) at Pine Island") %>%
  kable_styling(fixed_thead = T) %>% 
  column_spec(3, border_left = T)
  
  
```

### Chlorophyll *a*

```{r}
#| label: fig-chla-boxplots
#| echo: false
#| fig.cap: Historical monthly chlorophyll a at each SWMP water quality station with current year monthly value overlaid in red. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = chla_n, 
                     threshold = TRUE,
                     thres_val = 6.6) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
             x = "Mar",
             y = 20,
             size = 3,
             color = "blue",
             label = "State Threshold 6.6 (\U00B5g/L)")

boxplot_currentyear(station = "gtmssnut", 
                     param = chla_n, 
                     threshold = TRUE,
                     thres_val = 4.0) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "Mar",
           y = 16,
           size = 3,
           color = "blue",
           label = "State Threshold 4.0 (\U00B5g/L)")

boxplot_currentyear(station = "gtmfmnut", 
                    param = chla_n, 
                    threshold = TRUE,
                    thres_val = 5.5) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "Mar",
           y = 24,
           size = 3,
           color = "blue",
           label = "State Threshold 5.5 (\U00B5g/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = chla_n, 
                     threshold = TRUE,
                     thres_val = 4.3) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "Mar",
           y = 38,
           size = 3,
           color = "blue",
           label = "State Threshold 4.3 (\U00B5g/L)")
```


```{r}
#| label: fig-chla-agm
#| echo: false
#| fig.cap: Annual geometric means of monthly chlorophyll *a* at each SWMP water quality station. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530. Values exceeding the threshold are indicated in red. The 2022 value is based upon data to date and does not equal an entire year of data.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

agm(station = "gtmpinut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 6.6) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 8,
           size = 3,
           color = "blue",
           label = "State Threshold 6.6 (\U00B5g/L)")

agm(station = "gtmssnut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 4.0) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 6,
           size = 3,
           color = "blue",
           label = "State Threshold 4.0 (\U00B5g/L)")

agm(station = "gtmfmnut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 5.5) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 6,
           size = 3,
           color = "blue",
           label = "State Threshold 5.5 (\U00B5g/L)")

agm(station = "gtmpcnut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 4.3) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 16,
           size = 3,
           color = "blue",
           label = "State Threshold 4.3 (\U00B5g/L)")

```

At the conclusion of 2023, variability and patterns in the monthly chlorophyll *a* data will be examined by decomposing the timeseries at each station using Equation 4 from @cloern2010 into an annual effect, mean seasonal pattern, and residual "events". 

::: callout-note
#### Equation 4 from @cloern2010:

$$
c_{ij}=Cy_im_j\epsilon_{ij}
$$ Where $c_{ij}$ is chlorophyll concentration in year $i$($i=1,...,N$) and month $j$($j=1,...12$); *C* is the long-term mean of the series; $y_i$ is the annual effect in the *i*th year; $m_j$ is the seasonal (monthly) effect in the *j*th month; and $\epsilon_{ij}$ is the residual.

-   If $y_i$ \>1, indicates a year with above-average mean Chl-*a*.
-   If $m_j$ \>1, indicates a mean monthly Chl-*a* greater than the annual mean.
-   If $\epsilon_{ij}$ \>1, indicates an observation greater than the expected value for that month and year.
:::

### Nitrogen

#### Total Nitrogen

TN is calculated by adding TKN and NO23.

Several of the NO23 values fell below the Florida Department of Environmental Protection Central Lab's minimum detection limit and therefore TN was not able to be calculated and is missing for quite a few months at all the stations (@fig-tn-boxplots). This was especially true at Fort Matanzas which has the least amount of TN data (@fig-tn-boxplots-3).

```{r}
#| label: fig-tn-boxplots
#| echo: false
#| fig.cap: Historical monthly total nitrogen at each SWMP water quality station with current year monthly value overlaid in red. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = tn,
                     threshold = TRUE,
                     thres_val = 0.65) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.2,
           size = 3,
           color = "blue",
           label = "State Threshold 0.65 (mg/L)")

boxplot_currentyear(station = "gtmssnut", 
                    param = tn,
                    threshold = TRUE,
                    thres_val = 0.55) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.0,
           size = 3,
           color = "blue",
           label = "State Threshold 0.55 (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = tn, 
                     threshold = TRUE,
                     thres_val = 0.55) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.0,
           size = 3,
           color = "blue",
           label = "State Threshold 0.55 (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = tn, 
                     threshold = TRUE,
                     thres_val = 1.10) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.7,
           size = 3,
           color = "blue",
           label = "State Threshold 1.10 (mg/L)")
```

#### Dissolved Inorganic Nitrogen

DIN is calculated by adding NO23 and ammonia (NH4). Due to the low or undetectable levels of NO23, DIN is also missing quite a bit of information (@fig-din-boxplots).

```{r}
#| label: fig-din-boxplots
#| echo: false
#| fig.cap: Historical monthly dissolved inorganic nitrogen at each SWMP water quality station with current year monthly value overlaid in red. 
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)") 

boxplot_currentyear(station = "gtmssnut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)")


```

#### Total Kjeldahl Nitrogen

TKN is the total concentration of organic nitrogen and ammonia. It is directly measured in our samples. 

```{r}
#| label: fig-tkn-boxplots
#| echo: false
#| fig.cap: Historical monthly total kjeldahl nitrogen at each SWMP water quality station with current year monthly value overlaid in red. 
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)") 

boxplot_currentyear(station = "gtmssnut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)")


```

### Phosphorus

#### Total Phosphorus

```{r}
#| label: fig-tp-boxplots
#| echo: false
#| fig.cap: Historical monthly total phosphorus at each SWMP water quality station with current year monthly value overlaid in red. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                    param = tp, 
                    threshold = TRUE,
                    thres_val = 0.105) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.15,
           size = 3,
           color = "blue",
           label = "State Threshold 0.105 (mg/L)")

boxplot_currentyear(station = "gtmssnut",
                    param = tp,
                    threshold = TRUE, 
                    thres_val = 0.11) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.12,
           size = 3,
           color = "blue",
           label = "State Threshold 0.11 (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = tp,
                     threshold = TRUE, 
                     thres_val = 0.111) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.12,
           size = 3,
           color = "blue",
           label = "State Threshold 0.111 (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = tp, 
                     threshold = TRUE,
                     thres_val = 0.123) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.25,
           size = 3,
           color = "blue",
           label = "State Threshold 0.123 (mg/L)")
```

#### Orthophosphate

```{r}
#| label: fig-po4-boxplots
#| echo: false
#| fig.cap: Historical monthly orthophosphate at each SWMP water quality station with current year monthly value overlaid in red.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 

boxplot_currentyear(station = "gtmssnut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 

boxplot_currentyear(station = "gtmfmnut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 

boxplot_currentyear(station = "gtmpcnut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 
```
