---
title: "Nutrient Indicators"
execute: 
  warning: false
  error: false
link-external-icon: true
link-external-newwindow: true
filters: 
  - lightbox
lightbox: auto
bibliography: references.bib
---

```{r}
#| label: data-load
#| include: false
library(here)

source(here::here('R', '00_loadpackages.R'))
source(here::here('R', '02.1_load_wrangle_NUT.R'))
source(here::here('R', '02.2_load_wrangle_WQ-and-MET.R'))
source(here::here('R', '03_graph_fxns.R'))
```

The State Thresholds used in the chlorophyll *a*, total nitrogen, and total phosphorus calculations come from Florida State Surface Water Quality Standards within the Florida Administrative Code.

::: callout-note
# Access the Standards

By visiting [Florida Rules](https://www.flrules.org/gateway/ruleno.asp?id=62-302.530)
:::

## Dissolved Oxygen

All stations experience seasonal patterns in dissolved oxygen with lower values in the warm, summer months (@fig-do). A distinctive jump in dissolved oxygen occurred at all sites during the storm period of Hurricane Ian (September 28 - October 1, 2022) as well as Hurricane Nicole (November 9-11, 2022).

```{r}
#| label: fig-do
#| echo: false
#| fig.cap: Historical daily range of dissolved oxygen at each SWMP water quality station with current year daily average overlaid in blue. Horizontal dashed red line indicates 2mg/L hypoxia threshold.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

pi |> 
  filter(do_mgl > 0) |> 
historical_daily_range(param = 'do_mgl', target_yr = 2022, criteria = 2)

historical_daily_range(ss, param = 'do_mgl', target_yr = 2022, criteria = 2)

historical_daily_range(filter(fm, do_mgl > 0), param = 'do_mgl', target_yr = 2022, criteria = 2)

pc |>
  filter(do_mgl < 20 & do_mgl > 0) |> 
  historical_daily_range(param = 'do_mgl', target_yr = 2022, criteria = 2)

```

Pellicer Creek is located within a waterbody with known impairments for dissolved oxygen, chlorophyll *a*, and fecal coliform bacteria. As such, this station frequently experiences low dissolved oxygen values, particularly in the summer months; this year has been no exception (@fig-do-pc). In fact, the dissolved oxygen in late August was lower than values observed during the time of Hurricane Irma (@fig-do-pc-1). Values began to rise in September, but it is likely that the storm surge associated with Hurricane Ian in late September helped to bring in and further oxygenate this system (@fig-do-pc-2).

```{r}
#| label: fig-do-pc
#| echo: false
#| fig.cap: Continuous dissolved oxygen at Pellicer Creek with monthly average overlaid in solid gray line. Conditions are indicated as good when above 5 mg/L, fair between 2 and 5 mg/L, and poor if less than 2 mg/L.
#| fig.subcap:
#|     - "For the past five years"
#|     - "For the current 2022 year"
#| layout-ncol: 2
    
threshold_criteria_plot(pc, param = 'do_mgl', rng = c(2017, 2022), 
                        thresholds = c(2,5),
                        threshold_labs = c('Poor', 'Fair', 'Good'), 
                        monthly_smooth = TRUE, 
                        threshold_cols = c('#FEC596', '#FFFFCC', '#ABD9E9')) +
  theme(axis.text = element_text(color = "black", size = 18),
        axis.title.y = element_text(size = 20))



threshold_criteria_plot(pc, param = 'do_mgl', rng = c(2022, 2022), 
                        thresholds = c(2,5),
                        threshold_labs = c('Poor', 'Fair', 'Good'), 
                        monthly_smooth = TRUE, 
                        threshold_cols = c('#FEC596', '#FFFFCC', '#ABD9E9')) +
  theme(axis.text = element_text(color = "black", size = 18),
        axis.title.y = element_text(size = 20))

```

## Chlorophyll *a*

Elevated chlorophyll *a* levels were observed at the three intracoastal waterway sites over several months this year (@fig-chla-boxplots). The levels were especially high at Pine Island from April through August with each month seeing the highest value this year compared to all the data from previous years; the highest observed chlorophyll in the entire time series was observed in April (18.5 $\mu$g/L) (@fig-chla-boxplots-1). These extremely high monthly values resulted in an annual geometric mean (AGM) chlorophyll value above the state threshold for the second time in the 19 year data set at Pine Island (@fig-chla-agm-1).

```{r}
#| label: fig-chla-boxplots
#| echo: false
#| fig.cap: Historical monthly chlorophyll a at each SWMP water quality station with current year monthly value overlaid in red. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = chla_n, 
                     threshold = TRUE,
                     thres_val = 6.6) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
             x = "Mar",
             y = 20,
             size = 3,
             color = "blue",
             label = "State Threshold 6.6 (\U00B5g/L)")

boxplot_currentyear(station = "gtmssnut", 
                     param = chla_n, 
                     threshold = TRUE,
                     thres_val = 4.0) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "Mar",
           y = 16,
           size = 3,
           color = "blue",
           label = "State Threshold 4.0 (\U00B5g/L)")

boxplot_currentyear(station = "gtmfmnut", 
                    param = chla_n, 
                    threshold = TRUE,
                    thres_val = 5.5) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "Mar",
           y = 24,
           size = 3,
           color = "blue",
           label = "State Threshold 5.5 (\U00B5g/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = chla_n, 
                     threshold = TRUE,
                     thres_val = 4.3) +
  labs(y = "Mean Monthly Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "Mar",
           y = 38,
           size = 3,
           color = "blue",
           label = "State Threshold 4.3 (\U00B5g/L)")
```

San Sebastian also had the highest chlorophyll observed in January and February of this year compared to the same months in all previous years (6.85 and 5.6 $\mu$g/L, respectively; @fig-chla-boxplots-2). High chlorophyll *a* at all stations in January are likely due to uncharacteristically high temperatures that month (@fig-range-temp). San Sebastian and Pellicer Creek continued to see high chlorophyll at those sites (@fig-chla-agm-2; @fig-chla-agm-4).

Interestingly, though much higher seasonal chlorophyll was observed at Fort Matanzas in the first quarter, August was much lower than average (@fig-chla-boxplots-3). Fort Matanzas also did not have high annual mean chlorophyll (@fig-chla-agm-3). It continues to be a station with observed regularly with low chlorophyll *a*. Values were much lower (some even below threshold levels) at all sites in the last quarter of the year.

```{r}
#| label: fig-chla-agm
#| echo: false
#| fig.cap: Annual geometric means of monthly chlorophyll *a* at each SWMP water quality station. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530. Values exceeding the threshold are indicated in red. The 2022 value is based upon data to date and does not equal an entire year of data.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

agm(station = "gtmpinut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 6.6) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 8,
           size = 3,
           color = "blue",
           label = "State Threshold 6.6 (\U00B5g/L)")

agm(station = "gtmssnut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 4.0) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 6,
           size = 3,
           color = "blue",
           label = "State Threshold 4.0 (\U00B5g/L)")

agm(station = "gtmfmnut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 5.5) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 6,
           size = 3,
           color = "blue",
           label = "State Threshold 5.5 (\U00B5g/L)")

agm(station = "gtmpcnut", 
    param = CHLA_agm, 
    threshold = TRUE,
    thres_val = 4.3) +
  labs(y = "Geo.Mean Annual Chlorophyll-a (\U00B5g/L)") +
  annotate("text",
           x = "2006",
           y = 16,
           size = 3,
           color = "blue",
           label = "State Threshold 4.3 (\U00B5g/L)")

```

Using methods from @cloern2010, variability and patterns in the monthly chlorophyll *a* data were further extracted by decomposing the timeseries at each station using Equation 4 into an annual effect, mean seasonal pattern, and residual "events".

::: callout-note
### Equation 4 from @cloern2010:

$$
c_{ij}=Cy_im_j\epsilon_{ij}
$$ Where $c_{ij}$ is chlorophyll concentration in year $i$($i=1,...,N$) and month $j$($j=1,...12$); *C* is the long-term mean of the series; $y_i$ is the annual effect in the *i*th year; $m_j$ is the seasonal (monthly) effect in the *j*th month; and $\epsilon_{ij}$ is the residual.

-   If $y_i$ \>1, indicates a year with above-average mean Chl-*a*.
-   If $m_j$ \>1, indicates a mean monthly Chl-*a* greater than the annual mean.
-   If $\epsilon_{ij}$ \>1, indicates an observation greater than the expected value for that month and year.
:::

```{r}
#| label: decomp-fxn
#| include: false

# function comes from Cloern and Jassby 2010 Supplemental Material

decomp.mult <- function(x, startyr = NULL , endyr = NULL, event = T){
#R2 .8.0 3/2/09 4:48 PM
  if(class(x)[1]!='mts') dim(x)=c(length(x),1)
  
  if(is.null(startyr)) startyr <- start(x)[1]
  
  if(is.null(endyr)) endyr <- end(x)[1]
  
  d <- window(x, start = c(startyr,1), end = c(endyr,12), extend = T)
  
  results <- vector('list',dim(d)[2])
  
  names(results)= colnames(d)
  
  for(site in colnames (d)){
    d1=d[,site]
    
    #long-term mean
    grandmean = mean(d1, na.rm = T)
    
    # annual component
    annualmean = aggregate(d1,1, mean, na.rm = T)
    annualmeanreps = as.vector(t(matrix(rep(annualmean,12), 
                                        ncol = 12)))
    interann = ts(annualmeanreps,s=c(startyr,1),f = 12) / grandmean
    # remaining components
    if(event) {
      # monthly component
      d2 = matrix(d1, nrow = 12)
      monthdev = sweep(d2,2, annualmean ,'/')
      monthmean = apply(monthdev,1, mean , na.rm = T)
      season = ts(rep(monthmean, endyr - startyr + 1), 
                  s=c(startyr, 1), f = 12)
      # events component
      resids = sweep(monthdev, 1, monthmean , '/')
      events = ts(as.vector(resids),
                  s=c(startyr, 1),f = 12)
    }
    else {
      # monthly component
      season = d1/(grandmean * interann)
    }
    # prepare output
    if(event) dcomp = ts.union(d1, grandmean, interann, season,
      events) else dcomp =ts.union (d1, grandmean, interann,
      season)
    colnames(dcomp)[1]= 'original'
    results[[site]]= dcomp
    }
    if(class(x)[1]!='mts') results[[1]] else results
}
```

```{r}
#| label: dat-ts
#| include: false

# 2005-01-01 is missing from all four sites, need to build in a placeholder for it
mis <- tribble(
  ~datetimestamp, ~chla_n,
  "2005-01-01", NA
) %>% mutate(datetimestamp = as.Date(datetimestamp),
             chla_n = as.numeric(chla_n))

# create timeseries objects of chlorophyll data
chla_pi <- ts(as.data.frame(pi_nut_mo %>% 
                              filter(datetimestamp > "2002-12-31" &
                                       datetimestamp < "2022-01-01") %>%
                              select(datetimestamp, chla_n) %>% 
                              bind_rows(mis) %>% 
                              arrange(datetimestamp) %>% 
                              select(chla_n)),
              start = c(2003,1), # sampling started 2002-05, but 2003 is first complete year
              end = c(2022,12),
              frequency = 12) # monthly

chla_ss <- ts(as.data.frame(ss_nut_mo %>% 
                              filter(datetimestamp > "2002-12-31" &
                                       datetimestamp < "2022-01-01") %>%
                              select(datetimestamp, chla_n) %>% 
                              bind_rows(mis) %>% 
                              arrange(datetimestamp) %>% 
                              select(chla_n)),
              start = c(2003,1), # sampling started 2002-05, but 2003 is first complete year
              end = c(2022,12), # time period ended 2022, but 2021 is last complete year
              frequency = 12) # monthly

chla_fm <- ts(as.data.frame(fm_nut_mo %>% 
                              filter(datetimestamp > "2002-12-31" &
                                       datetimestamp < "2022-01-01") %>%
                              select(datetimestamp, chla_n) %>% 
                              bind_rows(mis) %>% 
                              arrange(datetimestamp) %>% 
                              select(chla_n)),
              start = c(2003,1), # sampling started 2002-05, but 2003 is first complete year
              end = c(2022,12), 
              frequency = 12) # monthly

chla_pc <- ts(as.data.frame(pc_nut_mo %>% 
                              filter(datetimestamp > "2002-12-31" &
                                       datetimestamp < "2022-01-01") %>%
                              select(datetimestamp, chla_n) %>% 
                              bind_rows(mis) %>% 
                              arrange(datetimestamp) %>% 
                              select(chla_n)),
              start = c(2003,1), # sampling started 2002-05, but 2003 is first complete year
              end = c(2022,12),
              frequency = 12) # monthly

rm(mis)
```

```{r}
#| label: decomp
#| include: false

decomp_pi <- data.frame(decomp.mult(chla_pi), 
                        date = as.Date(time(decomp.mult(chla_pi))))
summary(decomp.mult(chla_pi))

decomp_ss <- data.frame(decomp.mult(chla_ss), 
                        date = as.Date(time(decomp.mult(chla_ss))))
summary(decomp.mult(chla_ss))

decomp_fm <- data.frame(decomp.mult(chla_fm), 
                        date = as.Date(time(decomp.mult(chla_fm))))
summary(decomp.mult(chla_fm))

decomp_pc <- data.frame(decomp.mult(chla_pc), 
                        date = as.Date(time(decomp.mult(chla_pc))))
summary(decomp.mult(chla_pc))
```

```{r}
#| label: plot-fxn
#| include: false


mult_fig3 <- function(site){
  if(site == "PI") {
    a <- ggplot(decomp_pi) +
            geom_line(aes(x = date, y = original), color = "blue") +
            scale_x_date(date_breaks = "year", date_labels = "%Y") +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, "black", NA, NA)),
                  panel.grid.minor = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Chl-a \u00b5g/L',
                 title = "A. Pine Island")
    b <- decomp_pi %>% 
          mutate(year = year(date)) %>% 
          select(year, interann) %>% 
          unique() %>% 
          ggplot() +
            geom_col(aes(x = year, y = interann-1), fill = "red") +
            geom_hline(yintercept = 0, color = "red") +
            scale_x_continuous(breaks = seq(2000, 2022, by = 1)) +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, NA, "black", NA)),
                  panel.grid = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Annual')
    c <- ggplot(decomp_pi) +
          geom_col(aes(x = date, y = events-1), fill = "purple") +
          geom_hline(yintercept = 0, color = "purple") +
          scale_x_date(date_breaks = "year", date_labels = "%Y") +
          scale_y_continuous(limits = c(-1,2)) +
          theme_bw() +
          theme(axis.text.x = element_text(colour = c(NA, NA,"black", NA, NA)),
                panel.grid = element_blank(),
                axis.text = element_text(color = "black")) +
          labs(x = '',
               y = 'Residual')
    d <- decomp_pi %>% 
          mutate(month = month(date, label = T)) %>% 
          select(month, season) %>% 
          unique %>% 
          ggplot() +
            geom_col(aes(x = month, y = season-1), fill = "green") +
            geom_hline(yintercept = 0, color = "green") +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text = element_text(color = "black"),
                  panel.grid = element_blank()) +
            labs(x = '',
                 y = 'Season')
    a/b/c/d
  } else if(site == "SS") {
     a <- ggplot(decomp_ss) +
            geom_line(aes(x = date, y = original), color = "blue") +
            scale_x_date(date_breaks = "year", date_labels = "%Y") +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, "black", NA, NA)),
                  panel.grid.minor = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Chl-a \u00b5g/L',
                 title = "B. San Sebastian")
    b <- decomp_ss %>% 
          mutate(year = year(date)) %>% 
          select(year, interann) %>% 
          unique() %>% 
          ggplot() +
            geom_col(aes(x = year, y = interann-1), fill = "red") +
            geom_hline(yintercept = 0, color = "red") +
            scale_x_continuous(breaks = seq(2000, 2022, by = 1)) +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, NA, "black", NA)),
                  panel.grid = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Annual')
    c <- ggplot(decomp_ss) +
          geom_col(aes(x = date, y = events-1), fill = "purple") +
          geom_hline(yintercept = 0, color = "purple") +
          scale_x_date(date_breaks = "year", date_labels = "%Y") +
          scale_y_continuous(limits = c(-1,2)) +
          theme_bw() +
          theme(axis.text.x = element_text(colour = c(NA, NA,"black", NA, NA)),
                panel.grid = element_blank(),
                axis.text = element_text(color = "black")) +
          labs(x = '',
               y = 'Residual')
    d <- decomp_ss %>% 
          mutate(month = month(date, label = T)) %>% 
          select(month, season) %>% 
          unique %>% 
          ggplot() +
            geom_col(aes(x = month, y = season-1), fill = "green") +
            geom_hline(yintercept = 0, color = "green") +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text = element_text(color = "black"),
                  panel.grid = element_blank()) +
            labs(x = '',
                 y = 'Season')
    a/b/c/d
  } else if(site == "FM"){
    a <- ggplot(decomp_fm) +
            geom_line(aes(x = date, y = original), color = "blue") +
            scale_x_date(date_breaks = "year", date_labels = "%Y") +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, "black", NA, NA)),
                  panel.grid.minor = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Chl-a \u00b5g/L',
                 title = "C. Fort Matanzas")
    b <- decomp_fm %>% 
          mutate(year = year(date)) %>% 
          select(year, interann) %>% 
          unique() %>% 
          ggplot() +
            geom_col(aes(x = year, y = interann-1), fill = "red") +
            geom_hline(yintercept = 0, color = "red") +
            scale_x_continuous(breaks = seq(2000, 2022, by = 1)) +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, NA, "black", NA)),
                  panel.grid = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Annual')
    c <- ggplot(decomp_fm) +
          geom_col(aes(x = date, y = events-1), fill = "purple") +
          geom_hline(yintercept = 0, color = "purple") +
          scale_x_date(date_breaks = "year", date_labels = "%Y") +
          scale_y_continuous(limits = c(-1,2)) +
          theme_bw() +
          theme(axis.text.x = element_text(colour = c(NA, NA,"black", NA, NA)),
                panel.grid = element_blank(),
                axis.text = element_text(color = "black")) +
          labs(x = '',
               y = 'Residual')
    d <- decomp_fm %>% 
          mutate(month = month(date, label = T)) %>% 
          select(month, season) %>% 
          unique %>% 
          ggplot() +
            geom_col(aes(x = month, y = season-1), fill = "green") +
            geom_hline(yintercept = 0, color = "green") +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text = element_text(color = "black"),
                  panel.grid = element_blank()) +
            labs(x = '',
                 y = 'Season')
    a/b/c/d
  } else {
    a <- ggplot(decomp_pc) +
            geom_line(aes(x = date, y = original), color = "blue") +
            scale_x_date(date_breaks = "year", date_labels = "%Y") +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, "black", NA, NA)),
                  panel.grid.minor = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Chl-a \u00b5g/L',
                 title = "D. Pellicer Creek")
    b <- decomp_pc %>% 
          mutate(year = year(date)) %>% 
          select(year, interann) %>% 
          unique() %>% 
          ggplot() +
            geom_col(aes(x = year, y = interann-1), fill = "red") +
            geom_hline(yintercept = 0, color = "red") +
            scale_x_continuous(breaks = seq(2000, 2022, by = 1)) +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text.x = element_text(colour = c(NA, NA, NA, "black", NA)),
                  panel.grid = element_blank(),
                  axis.text = element_text(color = "black")) +
            labs(x = '',
                 y = 'Annual')
    c <- ggplot(decomp_pc) +
          geom_col(aes(x = date, y = events-1), fill = "purple") +
          geom_hline(yintercept = 0, color = "purple") +
          scale_x_date(date_breaks = "year", date_labels = "%Y") +
          scale_y_continuous(limits = c(-1,2)) +
          theme_bw() +
          theme(axis.text.x = element_text(colour = c(NA, NA,"black", NA, NA)),
                panel.grid = element_blank(),
                axis.text = element_text(color = "black")) +
          labs(x = '',
               y = 'Residual')
    d <- decomp_pc %>% 
          mutate(month = month(date, label = T)) %>% 
          select(month, season) %>% 
          unique %>% 
          ggplot() +
            geom_col(aes(x = month, y = season-1), fill = "green") +
            geom_hline(yintercept = 0, color = "green") +
            scale_y_continuous(limits = c(-1,1)) +
            theme_bw() +
            theme(axis.text = element_text(color = "black"),
                  panel.grid = element_blank()) +
            labs(x = '',
                 y = 'Season')
    a/b/c/d
  }
}

```

Based on these methods, Pine Island has observed higher annual chlorophyll *a* in the past three years. Additionally, the five months in 2022 in which the highest monthly values of chlorophyll *a* were recorded were identified in the residual "events" that were high, they were not the highest deviations in the timeseries; that value occurred in 2015 (@fig-chla-decomp-1). All other stations observed lower annual variability in 2022 and each station shows a distinctive seasonal pattern with the highest values occurring from April through August, particularly at Pine Island and Pellicer Creek (@fig-chla-decomp). San Sebastian and Fort Matanzas, likely due to their proximity to inlets, appear to have a more protracted season which further extends into October (@fig-chla-decomp-2; @fig-chla-decomp-3). 

```{r}
#| label: fig-chla-decomp
#| echo: false
#| fig-cap: Phytoplankton variability at each of the four SWMP water quality stations. Blue lines are monthly series of Chl-*a* concentration; red bars are annual components $y$ showing deviation from 1; purple lines are residual components $\epsilon$ showing deviation from 1, and green bars are standard seasonal patterns $m$ showing deviations from 1.
#| fig-subcap: 
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"

mult_fig3(site = "PI")

mult_fig3(site = "SS")

mult_fig3(site = "FM")

mult_fig3(site = "PC")

```

## Nitrogen

Interpretations of patterns in total nitrogen (TN) and dissolved inorganic nitrogen (DIN) are difficult this year since quite a bit of the values were not able to be calculated (see below). However, there is a pattern in total kjeldahl nitrogen (TKN) at all sites (@fig-tkn-boxplots). At Pine Island, there are very low values between January-March and elevated levels from April-July (@fig-tkn-boxplots-1). This distinctive shift between March and April may likely be in response to the large amounts of rainfall the system experienced in March (@fig-rainfall-2). This response was seen at all the other sites, especially Pellicer Creek, which is also a more brackish site and sensitive to upland freshwater deposition (@fig-tkn-boxplots-2).

There was enough data in the last quarter of the year at Pine Island to show high TN values in September, which exceeded the threshold, though fell within normal ranges for that month (@fig-tn-boxplots-1). High TN in April at Pellicer Creek, coupled with the high chlorophyll *a* values in this same time, could be a consequence of the large rain the area experienced during March (@fig-rainfall-2; @fig-tn-boxplots-4). 

All DIN above detection limit were reported less than 0.1 (mg/L) at all sites for all months this year. The higher values appear to be in months following large rain events such as October at Pellicer Creek (@fig-din-boxplots-4) and in April at San Sebastian (@fig-din-boxplots-2) and Fort Matanzas (@fig-din-boxplots-3).

### Total Nitrogen

TN is a calculated value based on TKN and NO23.

This year, several of the NO23 values fell below the Florida Department of Environmental Protection Central Lab's minimum detection limit and therefore TN was not able to be calculated and is missing for quite a few months at all the stations (@fig-tn-boxplots). This was especially true at Fort Matanzas which has the least amount of TN data (@fig-tn-boxplots-3).

```{r}
#| label: fig-tn-boxplots
#| echo: false
#| fig.cap: Historical monthly total nitrogen at each SWMP water quality station with current year monthly value overlaid in red. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = tn,
                     threshold = TRUE,
                     thres_val = 0.65) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.2,
           size = 3,
           color = "blue",
           label = "State Threshold 0.65 (mg/L)")

boxplot_currentyear(station = "gtmssnut", 
                    param = tn,
                    threshold = TRUE,
                    thres_val = 0.55) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.0,
           size = 3,
           color = "blue",
           label = "State Threshold 0.55 (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = tn, 
                     threshold = TRUE,
                     thres_val = 0.55) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.0,
           size = 3,
           color = "blue",
           label = "State Threshold 0.55 (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = tn, 
                     threshold = TRUE,
                     thres_val = 1.10) +
  labs(y = "Mean Monthly Total Nitrogen (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 1.7,
           size = 3,
           color = "blue",
           label = "State Threshold 1.10 (mg/L)")
```

### Dissolved Inorganic Nitrogen

DIN is a calculated value based on NO23 and ammonia (NH4). Due to the low or undetectable levels of NO23, DIN is also missing quite a bit of information for this year (@fig-din-boxplots). 

```{r}
#| label: fig-din-boxplots
#| echo: false
#| fig.cap: Historical monthly dissolved inorganic nitrogen at each SWMP water quality station with current year monthly value overlaid in red. 
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)") 

boxplot_currentyear(station = "gtmssnut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = din, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Dissolved Inorganic Nitrogen (mg/L)")


```

### Total Kjeldahl Nitrogen

TKN is the total concentration of organic nitrogen and ammonia. It is directly measured in our samples. Elevated levels of TKN were recorded in April at all sites (@fig-tkn-boxplots). These elevated values continued through August at Pine Island (@fig-tkn-boxplots-1). Pellicer Creek appears to have more variability in TKN month-to-month than the other stations (@fig-tkn-boxplots-4).

```{r}
#| label: fig-tkn-boxplots
#| echo: false
#| fig.cap: Historical monthly total kjeldahl nitrogen at each SWMP water quality station with current year monthly value overlaid in red. 
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)") 

boxplot_currentyear(station = "gtmssnut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = tkn, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Total Kjeldahl Nitrogen (mg/L)")


```

## Phosphorus

Total phosphorus continues to fall below the state threshold for every station (@fig-tp-boxplots). Though below the state threshold, Pine Island experienced elevated TP levels April-July (@fig-tp-boxplots-1). Total phosphorus levels were typically much lower than average at the remaining sites during that same time period. Pellicer Creek recorded an elevated TP level in August 2022, quite close to the state threshold value. Although higher than all the other values recorded in 2022, it remains in the average range for what is recorded for this month (@fig-tp-boxplots-4). In general, orthophosphate levels typically fell within the historical ranges at all sites and mirrored patterns observed in TP; however, elevated amounts were observed in April and August at all four stations (@fig-po4-boxplots).

### Total Phosphorus

```{r}
#| label: fig-tp-boxplots
#| echo: false
#| fig.cap: Historical monthly total phosphorus at each SWMP water quality station with current year monthly value overlaid in red. Horizontal dashed blue line indicates the state threshold, as found in F.A.C 62-302.530.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                    param = tp, 
                    threshold = TRUE,
                    thres_val = 0.105) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.15,
           size = 3,
           color = "blue",
           label = "State Threshold 0.105 (mg/L)")

boxplot_currentyear(station = "gtmssnut",
                    param = tp,
                    threshold = TRUE, 
                    thres_val = 0.11) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.12,
           size = 3,
           color = "blue",
           label = "State Threshold 0.11 (mg/L)")

boxplot_currentyear(station = "gtmfmnut", 
                     param = tp,
                     threshold = TRUE, 
                     thres_val = 0.111) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.12,
           size = 3,
           color = "blue",
           label = "State Threshold 0.111 (mg/L)")

boxplot_currentyear(station = "gtmpcnut", 
                     param = tp, 
                     threshold = TRUE,
                     thres_val = 0.123) +
  labs(y = "Mean Monthly Total Phosphorus (mg/L)") +
  annotate("text",
           x = "Mar",
           y = 0.25,
           size = 3,
           color = "blue",
           label = "State Threshold 0.123 (mg/L)")
```

### Orthophosphate

```{r}
#| label: fig-po4-boxplots
#| echo: false
#| fig.cap: Historical monthly orthophosphate at each SWMP water quality station with current year monthly value overlaid in red.
#| fig.subcap:
#|    - "Pine Island"
#|    - "San Sebastian"
#|    - "Fort Matanzas"
#|    - "Pellicer Creek"
#| layout-ncol: 2

boxplot_currentyear(station = "gtmpinut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 

boxplot_currentyear(station = "gtmssnut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 

boxplot_currentyear(station = "gtmfmnut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 

boxplot_currentyear(station = "gtmpcnut", 
                     param = po4f, 
                     threshold = FALSE) +
 labs(y = "Mean Monthly Orthophosphate (mg/L)") 
```
